Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-12 content/impl.tex
removing notion of set of functions with given signature; working on proofs

A micro-function $f\in\lcc\to \lcc$ defines how the map from receivers to types should
be updated when an instruction is executed.
The micro-function for most kinds of instructions is the identity.
On a call to and return from a specific method $m$ called on receiver $r$,
the micro-function restricts the receiver-to-type map to map $r$ only to
types consistent with the polymorphic dispatch to method $m$.
Finally, when an instruction assigns an object of unknown type to a receiver
$r$, the corresponding micro-function updates the map to map $r$ to the
set of all types. This is made precise by the following definition:
\begin{definition}\label{def:microfn}
Given a previously fixed set $S \subseteq R$ of receivers, the micro-function $\varepsilon_S(e)$
of a supergraph edge $e$ is defined as:
                    \otodo{define $n_1(e)$ globally somewhere early on, and consider using a more decriptive name like src instead of $n_1$}
  \begin{align}\label{eq:varepsilon}
      &\varepsilon_S(e)= \lambda m\,.\,\\
        &\begin{cases}
            m[r\to m(r)\cap \tau(s_\mathcal F,\,f)],
            &\text{\parbox{75mm}{
                if $e$ is a call-start edge, $r.c()$ is
                the call site at $\src e$, $f$ is the called
                procedure with signature $s_\mathcal F$,
                and $r\in S$;}}\\[3ex]
        \parbox{30mm}{\[\begin{array}{l}
                m[r\to m(r)\cap \tau(s_\mathcal F,\,f)]\\
            [v_1\to\bot_T] \ldots [v_k\to\bot_T],
        \end{array}
        \]
        }
        &
        \text{\parbox{75mm}{
                if $e$ is an end-return edge,
                $r.c()$ is the call
                corresponding to the return node
                at $\target e$,
                $f$ is the called method with
                signature $s_\mathcal F$,
                $v_1,\,\dots,\,v_k\in S$ are the local variables
                in $f$,
                and $r\in S$;}}\\
            m\left[r\to \bot_T\right],
            &\text{\parbox{75mm}{if $\src e$ contains an assignment
                    for $r\in S$;}}\\
            m
                &\text{otherwise.}
        \end{cases}
  \end{align}
\end{definition}

@@

\subsection{Converting IDE Results to IFDS Results}\label{sec:ideToIfds}

For each program point $n$, the result of an IFDS analysis gives
a set of facts $d$ that may be reached at $n$. The result of an IDE analysis
pairs each such fact~$d$ with a lattice element $\ell$. Formally,
for an IFDS problem $P$, the result $\resultIFDS$ has type
$N \to D$.\mtodo{Should it be $N^*$?} Similarly, for an IFDS problem $Q$, the result $\resultIDE$
has type $N \to D \times L$.

Recall that in the equivalence transformation lattice $L^\equiv$, $\bot$ means reachable
and $\top$ means unreachable. Therefore, a result $\rho$ of the equivalence IDE analysis
is converted to an IFDS result as follows:
$\backEq(\rho) = \lambda n.\{d\mid\rho(n) = (d,\top)\}$.

In the correlated-calls transformation lattice $\lcc$, a map that maps any receiver
to the empty set of possible types means that the corresponding data flow path is
infeasible. Therefore, a result $\rho$ of the correlated-calls IDE analysis is
converted to an IFDS result as follows:
$\backCC(\rho) = \lambda n.\{d\mid\rho(n) = (d,\ell), \forall r\,.\, \ell(r) \neq \top\}$.

>>>>

A micro-function $f\in\lcc\to \lcc$ defines how the map from receivers to types should
be updated when an instruction is executed.
The micro-function for most kinds of instructions is the identity.
On a call to and return from a specific method $m$ called on receiver $r$,
the micro-function restricts the receiver-to-type map to map $r$ only to
types consistent with the polymorphic dispatch to method $m$.
Finally, when an instruction assigns an object of unknown type to a receiver
$r$, the corresponding micro-function updates the map to map $r$ to the
set of all types. This is made precise by the following definition:
\begin{definition}\label{def:microfn}
Given a previously fixed set $S \subseteq R$ of receivers, the micro-function $\varepsilon_S(e)$
of a supergraph edge $e$ is defined as:
                    \otodo{define $n_1(e)$ globally somewhere early on, and consider using a more decriptive name like src instead of $n_1$}
  \begin{align}\label{eq:varepsilon}
      &\varepsilon_S(e)= \lambda m\,.\,\\
        &\begin{cases}
            m[r\to m(r)\cap \tau(s,\,f)],
            &\text{\parbox{75mm}{
                if $e$ is a call-start edge, $r.c()$ is
                the call site at $\src e$, $f$ is the called
                procedure with signature $s$,
                and $r\in S$;}}\\[3ex]
        \parbox{30mm}{\[\begin{array}{l}
                m[r\to m(r)\cap \tau(s,\,f)]\\
            [v_1\to\bot_T] \ldots [v_k\to\bot_T],
        \end{array}
        \]
        }
        &
        \text{\parbox{75mm}{
                if $e$ is an end-return edge,
                $r.c()$ is the call
                corresponding to the return node
                at $\target e$,
                $f$ is the called method with
                signature $s$,
                $v_1,\,\dots,\,v_k\in S$ are the local variables
                in $f$,
                and $r\in S$;}}\\
            m\left[r\to \bot_T\right],
            &\text{\parbox{75mm}{if $\src e$ contains an assignment
                    for $r\in S$;}}\\
            m
                &\text{otherwise.}
        \end{cases}
  \end{align}
\end{definition}

@@

\subsection{Converting IDE Results to IFDS Results}\label{sec:ideToIfds}\mtodo{I changed the formalisms in this section. Please see if you agree to these changes.}

For each program point $n$, the result of an IFDS analysis gives
a set of facts $d$ that may be reached at $n$. The result of an IDE analysis
pairs each such fact~$d$ with a lattice element $\ell$. Formally,
for an IFDS problem $P$, the result $\resultIFDS$ has type
$N^* \to D$.
Similarly, for an IDE problem $Q$, the result $\resultIDE(Q)$
has type $N \to (D \to L)$.

Recall that in the equivalence transformation lattice $L^\equiv$, $\bot$ means reachable
and $\top$ means unreachable.
Therefore, a result $\rho=\resultIDE(\transEq(P))$ of the equivalence IDE analysis
is converted to an IFDS result as follows:
\begin{equation}\label{eq:transEqDef}
  \backEq(\rho) = \lambda n.\{d\mid\rho(n)(d)\ne\top\}.
\end{equation}

In the correlated-calls transformation lattice $\lcc$, a map that maps any receiver
to the empty set of possible types means that the corresponding data flow path is
infeasible. Therefore, a result $\rho=\resultIDE(\transCC S(P))$ of the correlated-calls IDE analysis is
converted to an IFDS result as follows:
\begin{equation}\label{eq:transCcDef}
  \backCC(\rho) = \lambda n.\{d\mid\forall r\in S\,.\,\rho(n)(d)\neq\top_T\}.
\end{equation}

### 2014-12-12 content/related.tex
removing notion of set of functions with given signature; working on proofs

\section{Related Work}
\otodo{Should we remove first initials in this section?}
The IFDS algorithm is an instance of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.
\commentout{
Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.
}
IFDS has been used to encode a variety of data-flow problems such as typestate analysis%
\commentout{(determining which operations can be performed on an object at a given program point)}~\cite{naeem2008typestate,DBLP:conf/pldi/ZhangMGNY14} and shape analysis%
\commentout{(detecting errors and validating properties of programs at compile time)}%
~\cite{DBLP:conf/birthday/KreikerRRSWY13}. IFDS has been broadly used~\cite{DBLP:conf/pldi/ArztRFBBKTOM14,tripp2009taj} and extended~\cite{DBLP:conf/sigsoft/LerchHBM14} to solve taint-analysis problems.

The IFDS algorithm is implemented in two popular static-analysis frameworks for Java bytecode, the T.~J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

@@

Several extensions of the IFDS algorithm have been proposed to improve its precision,
efficiency, and expressiveness. N.~Naeem and O.~Lhot\'ak~\cite{naeem2010pei} propose
four extensions: two to improve efficiency on certain kinds of IFDS problems,
one to improve expressiveness to a wider range of IFDS problems, and one
to improve precision when applying the IFDS algorithm to programs in SSA form.
All of these extensions are orthogonal to and could be combined with the approach
that we present in this paper. In contrast to these techniques, our approach targets
analysis precision, not efficiency or expressiveness.

@@

Two other extensions target efficiency. E.\,Bodden et~al.~\cite{bodden2013spl} presents a framework
for applying IFDS analyses to software product lines. Their approach enables the analysis of
all possible products derived from a product line in a single analysis pass. Like our approach,
their approach transforms IFDS problems to IDE problems. The micro-functions keep track
of the possible program variations specified by the product line.
J.\,Rodriguez and O.\,Lhot\'ak evaluate a parallelized implementation of the IFDS algorithm using actors~\cite{rodriguez2010concurrent} that can take advantage of multiple processors.
%However, neither of those works is concerned with improving the precision of IFDS results.

@@

The idea of using correlated calls to remove infeasible paths in data-flow analyses of object-oriented programs was introduced by F.\,Tip~\cite{DBLP:journals/scp/Tip15}. The possibility of using IDE to achieve this is mentioned, but not elaborated upon. Our work presents a concrete solution to the problem and an implementation of that solution.

M.~Sridharan et~al.'s improvements to the precision of pointer analysis
for JavaScript programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. However, instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
(the JavaScript analogue of fields) of an object. Unlike in more static languages,
property names in JavaScript are first class values (strings). When a loop indexed by
property names updates the properties of objects, the loop body is extracted into a function
and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.

>>>>

\section{Related Work}
\otodo{Should we remove first initials in this section?}
The IFDS algorithm is an instance of the functional approach to data-flow analysis developed by Sharir and Pnueli~\cite{pnueli1981two}.
\commentout{
Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.
}
IFDS has been used to encode a variety of data-flow problems such as typestate analysis%
\commentout{(determining which operations can be performed on an object at a given program point)}~\cite{naeem2008typestate,DBLP:conf/pldi/ZhangMGNY14} and shape analysis%
\commentout{(detecting errors and validating properties of programs at compile time)}%
~\cite{DBLP:conf/birthday/KreikerRRSWY13}. IFDS has been broadly used~\cite{DBLP:conf/pldi/ArztRFBBKTOM14,tripp2009taj} and extended~\cite{DBLP:conf/sigsoft/LerchHBM14} to solve taint-analysis problems.

The IFDS algorithm is implemented in two popular static-analysis frameworks for Java bytecode, the T.J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

@@

Several extensions of the IFDS algorithm have been proposed to improve its precision,
efficiency, and expressiveness. Naeem and Lhot\'ak~\cite{naeem2010pei} propose
four extensions: two to improve efficiency on certain kinds of IFDS problems,
one to improve expressiveness to a wider range of IFDS problems, and one
to improve precision when applying the IFDS algorithm to programs in SSA form.
All of these extensions are orthogonal to and could be combined with the approach
that we present in this paper. In contrast to these techniques, our approach targets
analysis precision, not efficiency or expressiveness.

@@

Two other extensions target efficiency. Bodden et~al.~\cite{bodden2013spl} presents a framework
for applying IFDS analyses to software product lines. Their approach enables the analysis of
all possible products derived from a product line in a single analysis pass. Like our approach,
their approach transforms IFDS problems to IDE problems. The micro-functions keep track
of the possible program variations specified by the product line.
Rodriguez and Lhot\'ak evaluate a parallelized implementation of the IFDS algorithm using actors~\cite{rodriguez2010concurrent} that can take advantage of multiple processors.
%However, neither of those works is concerned with improving the precision of IFDS results.

@@

The idea of using correlated calls to remove infeasible paths in data-flow analyses of object-oriented programs was introduced by Tip~\cite{DBLP:journals/scp/Tip15}. The possibility of using IDE to achieve this is mentioned, but not elaborated upon. Our work presents a concrete solution to the problem and an implementation of that solution.

Sridharan et~al.'s improvements to the precision of pointer analysis
for JavaScript programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. However, instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
(the JavaScript analogue of fields) of an object. Unlike in more static languages,
property names in JavaScript are first class values (strings). When a loop indexed by
property names updates the properties of objects, the loop body is extracted into a function
and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.
